---
title: Working with shape files
author: Roberto Supe
date: '2020-10-14'
slug: []
categories:
  - R
tags:
  - BNU
  - learning R
  - plot
  - R Markdown
  - R studio
  - tutorial
  - landscape metrics
lastmod: '2020-10-14T10:56:35+08:00'
keywords: []
description: ''
comment: no
toc: 
autoCollapseToc: no
postMetaInFooter: no
hiddenFromHomePage: no
contentCopyright: no
reward: no
mathjax: no
mathjaxEnableSingleDollar: no
mathjaxEnableAutoNumber: no
hideHeaderAndFooter: no
flowchartDiagrams:
  enable: no
  options: ''
sequenceDiagrams:
  enable: no
  options: ''
---

Inside the The data folder you will find two sets of shape files. They are composed of several files and in order for them to work propertly keep them all.

First, we will transform those shape files to Raster files with the help of rgdal.

```{r}
#install.packages("rgdal",dependencies = TRUE) #only once 
library(rgdal) 
library(raster)
```

Read the shape files with `readOGR` and use plot to look at the map. 
```{r}
assigment_map <- readOGR("Data-shp/GLYlanduse.shp") #read your file
plot(assigment_map)
```

Currently the map does not have a set of CRS so we will assign them to it because it is required for further analysis.[This page](https://spatialreference.org/ref/) has a lot of projections and CRS for your maps. R will require the text inclided in the **Proj4**  option, make sure you find your location or a CRS that includes the whole world. 

```{r}
#Check CRS of the map
crs(assigment_map)
```

```{r}
#map coordinates missing so we will set them to this system
newcrs <- CRS("+proj=merc +a=6378137 +b=6378137 +lat_ts=0.0 +lon_0=0.0 +x_0=0.0 +y_0=0 +k=1.0 +units=m +nadgrids=@null +wktext  +no_defs") 
crs(assigment_map)<-newcrs #add crs to a file
crs(assigment_map)
```

## Create raster files

Set up a raster "template" to use in rasterize() so our shape file can be transformed.

```{r}
r <- raster(ncol=180, nrow=180)
extent(r) <- extent(assigment_map)
```

Rasterize the shapefile with `rasterize` make sure to have a target variable of a type double (numbers). The function `str` gives us the structure of any object. 

```{r}
str(assigment_map)
```

We can see that `NUM` our variable for the raster is a factor so we will need to change that before the raster of directly inside the function. We will do the last.

```{r}
raster_map<-rasterize(assigment_map, r,as.double(assigment_map$NUM))# Transform NUM
class(raster_map)
crs(raster_map)<-newcrs #add the crs
plot(raster_map)
```

Export your new map with `writeRaster`
```{r,eval=FALSE}
writeRaster(raster_map, "assigment.tif") #export the file
```

* Try to do the same with the second map `Landuse` 

#Activities

1. Check your maps
2. Calculate one type of LSM for each level or 6, 4 and 1 LSM for the landscape, class and patch level, respectively.
3. Draw three graphs: one boxplot graph for your 6 the landscapes metrics, one bar graph for the classes and one for the patch level metric.
4. Export your data and figures


* Try to use as fewer lines of code as possible, You can get all the results in 2 or 3 lines. Run both maps at the same time as instructed.
